Using Scagnostics to find Interesting Projections of Multivariate Data

### 2018-09-13 scagnostics.Rmd
Moved scagnostics definitions to appendix.
Hiding guided tour printout.

Scagnostics are scatterplot diagnostics. They were primarily developed to extract pairs of interesting variables from a large collection.
XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia).
Nine graph-theoretic measures have been defined to characterise scatterplots. They are calculated from a set of elementary building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST). For definiteness all variables are rescaled to a [0,1] intervall, such that all measures will take values between 0 and 1 as well.
The following measures have been defined:

@@

```{r findpipe, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Pipe 1000 dataset.", echo=FALSE}
set.seed(2018)
pipeResc <- rescale(pipe1000) #unselect 2 variables to make optimisation easier
pipeTour <- save_history(pipeResc,
                         guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
                                     cooling=1, alpha = 0.5, max.tries = 5000))
pipeTourFull <- as.list(interpolate(pipeTour))
pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
ggplot(pipeTourResMelt, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

@@

```{r refine}
pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
pipeTourFull2 <- as.list(interpolate(pipeTour2))
pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
ggplot(pipeTourResMelt2, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

```{r testpipe2}
iLast <- length(pipeTourFull)
fProj <- pipeTourFull[[iLast]]
dProj <- as.tibble(pipeResc %*% fProj)
ggplot(dProj, aes(V1,V2)) + geom_point()
```

### Finding sine waves
```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Sine 1000 dataset.", echo=FALSE}
set.seed(2018)
sineResc <- rescale(sin1000) #unselect 2 variables to make optimisation easier
sineTour <- save_history(sineResc,
                         guided_tour(splineIndex()))
sineTourFull <- as.list(interpolate(sineTour))
sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
ggplot(sineTourResMelt, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

@@

>>>>

Scagnostics are scatterplot diagnostics. They were primarily developed to extract pairs of interesting variables from a large collection.
XXX (Ursula to fill in based loosely on material at the German language wikipedia page - it would be good to translate this page and provide an English version on wikipedia).
Nine graph-theoretic measures have been defined to characterise scatterplots. They are calculated from a set of elementary building blocks: the convex hull, the alpha hull and the minimal spanning tree (MST). For definiteness all variables are rescaled to a [0,1] intervall, such that all measures will take values between 0 and 1 as well.

@@

```{r findpipe, results="hide"}
set.seed(2018)
pipeResc <- rescale(pipe1000) #unselect 2 variables to make optimisation easier
pipeTour <- save_history(pipeResc,
                         guided_tour(mineIndex("TIC"), search_f = tourr:::search_better,
                                     cooling=1, alpha = 0.5, max.tries = 5000))
pipeTourFull <- as.list(interpolate(pipeTour))
```

```{r pipeFirstRun,fig.width=6, fig.height=7, fig.cap="Guided tour optimising MIC index for Pipe 1000 dataset."}
pipeTourRes <- getProj(pipe1000, pipeTourFull, "Pipe", 1000)
pipeTourResMelt <- melt(pipeTourRes, id=c("t", "name", "size"))
ggplot(pipeTourResMelt, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

@@

```{r refine, results="hide"}
pipeTour2 <- save_history(pipeResc, guided_tour(mineIndex("TIC")), start = fProj)
pipeTourFull2 <- as.list(interpolate(pipeTour2))
pipeTourRes2 <- getProj(pipe1000, pipeTourFull2, "Pipe", 1000)
pipeTourResMelt2 <- melt(pipeTourRes2, id=c("t", "name", "size"))
ggplot(pipeTourResMelt2, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")

```

```{r testpipe2}

iLast <- length(pipeTourFull2)
fProj <- pipeTourFull2[[iLast]]
dProj <- as.tibble(pipeResc %*% fProj)
ggplot(dProj, aes(V1,V2)) + geom_point()
```

### Finding sine waves
```{r findsine, fig.width=6, fig.height=7, fig.cap="Guided tour optimising TIC index for Sine 1000 dataset.", results="hide"}
set.seed(2018)
sineResc <- rescale(sin1000) #unselect 2 variables to make optimisation easier
sineTour <- save_history(sineResc,
                         guided_tour(splineIndex()))
sineTourFull <- as.list(interpolate(sineTour))
sineTourRes <- getProj(sin1000, sineTourFull, "Sine", 1000)
sineTourResMelt <- melt(sineTourRes, id=c("t", "name", "size"))
ggplot(sineTourResMelt, aes(x=t, y=value)) +
  geom_line() + 
  facet_wrap(~variable, ncol=1, scales = "free_y")
```

@@

# Apendix
The following measures have been defined:
